---
title: "STAT-451"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'stat451.html'))})
output: 
  html_document:
    css: statCSSBackground.css
    theme: readable
---

# Introducton 

Statistics-4541 : Applied Statistics for Engineers and Scientist was taught Summer term 2021 at PSU by Subhash Kochar. The problems on this page are from Probability & Statistics for Engineering & Sciences 9th Edition by Jay L. Devore - Duxbury Publisher, and the work mine. 

### 1.10 : Consider the strength data for beams given in Example 1.2.

a. Construct a stem-and leaf display of the data. What appears to be a representative strength value? Do the observations appear to be highly concentrated about the representative value or rather spread out?

```{r}
beam <- c(5.9, 7.2, 7.3, 6.3, 8.1, 6.8, 7.0, 7.6, 6.8, 6.5, 7.0, 6.3, 7.9, 9.0,
         8.2, 8.7, 7.8, 9.7, 7.4, 7.7, 9.7, 7.8, 7.7, 11.6, 11.3, 11.8, 10.7)
stem(beam)
```

A representative strength value would be 7.7, as more observations are concentrated around this value than any other. 

b. Does the display appear to be reasonably symmetric about a representative value, or would you describe its shape in some other way? 

```{r, echo=F}
hist(beam, 
     col = topo.colors(7), 
     ylim = c(0,10), 
     xlab= "Flexural Strength (in MegaPascal)",
     main = "Flexural Strngth of Concreate Beams")
abline(v=mean(beam), col="orange", lwd = 6)
abline(v=median(beam), col="red", lwd = 8)
```

No, I would argue that the data has a slight positive skewed to the right. I say this because the higher frequency values (in blues) seem to be to the left of the representative value (indicated by the red line), which is to the left of the mean (indicated by the orange line).

c. Do there appear to be any outlying strength values? 

Yes there seems to be a small density of values outlying around 11. 

d. What proportion of strength observations in this sample exceed 10 MPa? 
```{r}
proportion <- mean(as.integer(beam > 10))
print(paste(round(proportion*100,2), "%"))
```

An unlikely 14.81% of values exceed 10 MPa. 

### 1.12 : The accompanying specific gravity values for various wood types used in construction appeared in the article “Bolted Connection Design Values Based on European Yield Model” (J. of Structural Engr., 1993: 2169-2186): 

```{r}
wood.g <- c(.31, .35, .36, .36, .37, .38, .40, .40, .40,
            .41, .41, .42, .42, .42, .42, .42, .43, .44,
            .45, .46, .46, .47, .48, .48, .48, .51, .54,
            .54, .55, .58, .62, .66, .66, .67, .68, .75)
```

### Construct a stem-and-leaf display using repeated stems (see the previous exercise), and comment on any interesting features of the display. 

```{r, echo=F}
stem(wood.g, scale = 0.1)
stem(wood.g, scale = 0.5)
stem(wood.g)
```

Looking at the three stem-and-leaf displays there are a few interesting features about the gravity for various wood types used in construction data that stands out. Dropping the last digit in the data, the first stem-and-leaf display shows us that there’s an even number of values that are >0.5 and 0.5that appear most dense around 0.4. The second and third stem-and-leaf displays show us that these values are most dense about 0.42, with a small blip around 6.6 and obvious outlier for value 0.75. 

### 1.16 The article cited in Example 1.2 also gave the accompanying strengths observations for cylinders: 

```{r}
beam2 <- c(6.1, 5.8, 7.8, 7.1, 7.2, 9.2, 6.6, 8.3, 7.0, 8.3,
           7.8, 8.1, 7.4, 8.5, 8.9, 9.8, 9.7, 14.1, 12.6, 11.2)
```

a. Construct a comparative stem-and-leaf display (see the previous exercise) of the beam and cylinder data and then answer the questions in parts (b)-(d) of Exercise 10 for the observations on cylinders. 

```{r, echo=F}
stem(beam2)
```

* Does the display appear to be reasonably symmetric about a representative value, or would you describe its shape in some other way? 

```{r, echo=F}
# find mode 
sort(table(beam2))
mode <- ((8.3-7.8)/2)+7.8

hist(beam2, 
     col = topo.colors(7), 
     ylim = c(0,6),
     xlim = c(0, 15),
     xlab= "Flexural Strength (in MegaPascal)",
     main = "Flexural Strngth of Concreate Beams")
abline(v=mean(beam2), col="orange", lwd = 6)
abline(v=median(beam2), col="red", lwd = 8)
abline(v=mode, col="yellow", lwd = 6)
```

This data also appears to have a slight positve skew, shown by mean (orange line) to the right of the median (red line) and mode (yellow line).

* Do there appear to be any outlying strength values? 


* What proportion of strength observations in this sample exceed 10 MPa?


